2.2: Random Variables and Distributions - Class Notes

### 2019-09-21 content/class/08-class.Rmd
Clean up class notes appendix 8
More cleaning of class notes 8 appendix

```{r}
# install.packages("mosaic") install for first use
library(mosaic) # load 
```

You can create custom mathematical functions using mosaic by defining an R `function()` with multiple arguments. As a simple example, make the function `f(x) = 2x+x^2` as follows:

@@

There are some notational requirements from `R` for making functions. Any coefficient in front of a variable (such as the 2 in `2x` must be explicitly multiplied by the variable, as in `10*x`.)

@@

## To Graph

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function) #<<
```

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function, color = "blue", size = 2, linetype = "dashed") #<<
```

There are some standard statistical distributions built into R:

@@

| Action/Type | Prefix |
|--------|--------|
| random draw | `r` |
| density (pdf) | `d` |  
| cdf    | `p`    |
| quantile (inverse cdf) | `q` |

@@

| Distribution | Name |
|--------|--------|
| normal | `norm` |
| uniform | `unif` |  
| t    | `t`    |
| binomial    | `binom`    |
| negative binomial    | `nbinom`    |
| hyper    | `hypergeometric`    |
| weibull    | `weibull`    |
| beta    | `beta`    |
| gamma    | `gamma`    |

@@

### Some common examples:

@@

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = -400:600))+
  aes(x = x)+
  stat_function(fun = dnorm, args = list(mean = 200, sd = 200), color = "blue", size = 2, linetype = "dashed") #<<
```

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed") #<<
```

@@

```{r, fig.retina=3}
# graph normal distribution and shade area between -2 and 2
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed")+
  stat_function(fun = dnorm, xlim = c(-2,2), geom = "area", fill = "green", alpha=0.5)#<<
```

>>>>

```{r, message = F, warning = F}
# install.packages("mosaic") install for first use
library(mosaic) # load 
```

## Creating Mathematical Functions

You can create custom mathematical functions using mosaic by defining an R `function()` with multiple arguments. As a simple example, make the function $f(x) = 10x-x^2$ as follows:

@@

There are some notational requirements from `R` for making functions. Any coefficient in front of a variable (such as the 10 in `10x` must be explicitly multiplied by the variable, as in `10*x`).

@@

## Graphing Mathematical Functions

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function) 
```

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = 1:10))+
  aes(x = x)+
  stat_function(fun = my_function, color = "blue", size = 2, linetype = "dashed") 
```

## Bultin Statistical Functions

There are some standard statistical distributions built into R. They require a combination of a specific prefix and a distribution.

@@

| Action/Type | Prefix |
|--------|--------|
| random draw | `r` |
| density (pdf) | `d` |  
| cumulative density (cdf)    | `p`    |
| quantile (inverse cdf) | `q` |

@@

| Distribution | Name in R |
|--------|--------|
| Normal | `norm` |
| Uniform | `unif` |  
| Student's t    | `t`    |
| Binomial    | `binom`    |
| Negative binomial    | `nbinom`    |
| Hypergeometric    | `hyper`    |
| Weibull    | `weibull`    |
| Beta    | `beta`    |
| Gamma    | `gamma`    |

@@

## Some common examples:

@@

## Graphing Statistical Functions

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = -400:600))+
  aes(x = x)+
  stat_function(fun = dnorm, args = list(mean = 200, sd = 200), color = "blue", size = 2, linetype = "dashed") 
```

@@

```{r, fig.retina=3}
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed") 
```

@@

```{r, fig.retina=3}
# graph normal distribution and shade area between -2 and 2
ggplot(data = data.frame(x = -4:4))+
  aes(x = x)+
  stat_function(fun = dnorm, color = "blue", size = 2, linetype = "dashed")+
  stat_function(fun = dnorm, xlim = c(-2,2), geom = "area", fill = "green", alpha=0.5)
```
